Looking for a Signal in the Noise: Revisiting Obesity and the Microbiome

### 2016-05-23 submission/Sze_Obesity_PeerJ_2016.Rmd
add Pat data pt2 more readable code

**Results:** When we considered each dataset in isolation, there were no significant differences (P > 0.05) in B/F between non-obese and obese individuals. A total of `r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric != "bacteroidetes"), ]))` diversity measurements  were found to be significantly different between non-obese and obese indivdiuals (P < 0.05) (Shannon Diversity (n=`r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric == "shannon"), ]))`), observed richness (n=`r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric == "sobs"), ]))`), and evenness (n=`r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric == "shannoneven"), ]))`)). With individual studies, random forest was effective at identifying features within the datasets for differentiating between non-obese and obese individuals (mean AUC: `r format(round(mean(rand_forest.otu.summary$auc), 2), nsmall=2)`; s.d.=`r format(round(sd(AUCRFData$OOBAUCAll), 2), nsmall=2)`) and there was no significant difference between the studies (bonferroni adjusted P-value > 0.05). In 7/8 of the studies an OTU related to the Ruminococcaceae family was important in classification; however, the direction of the association varied between studies. When we pooled the results across studies, there was no significant difference in RR of obesity based on Shannon diversity (RR = `r format(round(10^test3R$LogFC[1],2), nsmall=2)`, CI = `r format(round(10^ShannonRes$ci.lb, 2), nsmall=2)`-`r format(round(10^ShannonRes$ci.ub, 2), nsmall=2)`, P-value = `r format(round(RRSDPvalue$QMp, 2), nsmall=2)`) and B/F ratio (RR = `r format(round(10^test3BF$LogFC[1],2), nsmall=2)`, CI = `r format(round(10^BFRes$ci.lb, 2), nsmall=2)`-`r format(round(10^BFRes$ci.ub, 2), nsmall=2)`, P-value = `r format(round(RRBFPvalue$QMp, 2), nsmall=2)`). When we normalized the Shannon diversity data and pooled studies there was also no significant difference between non-obese and obese individuals (difference in mean z-score = `r format(round(combZHTtest$estimate[1]-combZHTtest$estimate[2], 2), nsmall=2)`; P = `r format(round(combZHTtest$p.value, 2), nsmall=2)`). There was also no significant difference in the normalized B/F ratio between non-obese and obese individuals (difference in mean z-score = `r format(round(combZLogBFTtest$estimate[1]-combZLogBFTtest$estimate[2], 2), nsmall=2)`; P = `r format(round(combZLogBFTtest$p.value, 2), nsmall=2)`).  Finally, when completing both power and sample size calculations and simulations we found that most of the studies examined were not sufficiently powered to detect differences in any of the measurements used.

>>>>

**Results:** When we considered each dataset in isolation, there were no significant differences (P > 0.05) in B/F between non-obese and obese individuals. A total of `r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric != "bacteroidetes"), ]))` diversity measurements  were found to be significantly different between non-obese and obese indivdiuals (P < 0.05) (Shannon Diversity (n=`r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric == "shannon"), ]))`), observed richness (n=`r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric == "sobs"), ]))`), and evenness (n=`r length(rownames(alpha_tests.summary[which(alpha_tests.summary$p.value < 0.05 & alpha_tests.summary$metric == "shannoneven"), ]))`)). With individual studies, random forest was effective at identifying features within the datasets for differentiating between non-obese and obese individuals for both the genus (mean AUC: `r format(round(mean(rand_forest.genus.summary$auc), 2), nsmall=2)`; s.d.=`r format(round(sd(rand_forest.genus.summary$auc), 2), nsmall=2)`) and OTU (mean AUC: `r format(round(mean(rand_forest.otu.summary$auc), 2), nsmall=2)`; s.d.=`r format(round(sd(rand_forest.otu.summary$auc), 2), nsmall=2)`) level, and there was no significant difference between the studies (bonferroni adjusted P-value > 0.05). When we pooled the results across studies, there was a significant difference in RR of obesity based on Shannon diversity (RR = `r format(round(RR_composite['shannon', 'rr'],2), nsmall=2)`, CI = `r format(round(RR_composite['shannon', 'ci_lb'], 2), nsmall=2)`-`r format(round(RR_composite['shannon', 'ci_ub'], 2), nsmall=2)`, P-value = `r format(round(RR_composite['shannon', 'p_value'], 3), nsmall=3)`) but not for B/F ratio (RR = `r format(round(RR_composite['bf_ratio', 'rr'],2), nsmall=2)`, CI = `r format(round(RR_composite['bf_ratio', 'ci_lb'], 2), nsmall=2)`-`r format(round(RR_composite['bf_ratio', 'ci_ub'], 2), nsmall=2)`, P-value = `r format(round(RR_composite['bf_ratio', 'p_value'], 3), nsmall=3)`). When we normalized the Shannon diversity data and pooled studies there was also no significant difference between non-obese and obese individuals (difference in mean z-score = `r format(round(combZHTtest$estimate[1]-combZHTtest$estimate[2], 2), nsmall=2)`; P = `r format(round(combZHTtest$p.value, 2), nsmall=2)`). There was also no significant difference in the normalized B/F ratio between non-obese and obese individuals (difference in mean z-score = `r format(round(combZLogBFTtest$estimate[1]-combZLogBFTtest$estimate[2], 2), nsmall=2)`; P = `r format(round(combZLogBFTtest$p.value, 2), nsmall=2)`).  Finally, when completing both power and sample size calculations and simulations we found that most of the studies examined were not sufficiently powered to detect differences in any of the measurements used.
